---
title: "Western Atlantic mackerel adults parameters for DEPM"
author: "mkorta modified from DEPM_mackerel_script of PSanpedro"
date: "4/7/2023"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    theme: flatly
    #code_folding: hide
    number_sections: true
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE, fig.width=12, fig.height=8)

knitr::opts_chunk$set(
 fig.width = 6,
 fig.asp = 0.8,
 out.width = "80%"
)

```
*Lola Garabana. Presentation WGMEGS meeting,April 17-21 Madrid.*

```{r echo=F, message=FALSE, warning=FALSE}
# Load libraries 
#library(ggmap)
library(ggplot2)
library(MASS) # masks "select" in dplyr
library(janitor)
library(dplyr)
library(tidyverse)
library(kableExtra)
library(SDMTools)
library(broom)
library(sf)
library(rworldmap)
library(rworldxtra)
```

```{r echo=F, message=FALSE, warning=FALSE}

# ---------------- Data loading and checks ------------------- #

# dat
adults.dat <- read.csv("dat/Datos DEPM 2022.csv")

# map
world <- getMap(resolution = "high")
world <- st_as_sf(world)
```

```{r echo=F, message=FALSE, warning=FALSE}
# Exploratory of data
# dim(adults.dat)
# head(adults.dat)  
# str(adults.dat)

# Identify duplicate records
#anyDuplicated(adults.dat)

# Remove duplicate 
# if(data.dup != 0) {
#   adults.dat <- adults.dat[!duplicated(adults.dat), ]
#   }
  
# check records without ...  
# adults.dat[is.na(adults.dat$Haul),]  
# adults.dat[is.na(adults.dat$Sampling),]  
# adults.dat[is.na(adults.dat$Period),]  

# Validate format for some variables 
adults.dat$Haul <- as.factor(as.character(adults.dat$Haul))
adults.dat$Sampling <- as.factor(as.character(adults.dat$Sampling))
adults.dat$Period <- as.factor(as.character(adults.dat$Period))



# Change "" levels for some factor variables by NA:
#levels(adults.dat$Oocyte)[1] <- NA
#levels(adults.dat$Age)[1] <- NA
```

```{r echo=F, message=FALSE, warning=FALSE}

# ---------------- Define values  ------------------- #

# Minimum maturity stage to be adult (for males and females)
stage.mat <- 2 

# Minimum samples 
min.totnumb <- 20 # minimum sampling number by haul to do calculations  

# Subsample weight
Wcent <- 0.1

```

# Raw Data Summary {.tabset .tabset-tabs}

## Tables

**By period**


```{r echo=F, message=FALSE, warning=FALSE}
nHaul <- adults.dat %>% group_by(Period) %>%  summarise(nHaul= length(unique(Haul)))

nfish <- adults.dat %>% group_by (Period, Sampling) %>%
         summarise(nFish=n())%>%
         pivot_wider(names_from = Sampling, values_from = nFish)%>%
         replace(is.na(.), 0)%>%
         rename("R_nFish" = "R", "D_nFish" = "D")%>%
         mutate(nFish = sum(c_across(where(is.numeric))))

# Females in Mat>=stage.mat 
nMatFem <- adults.dat %>% group_by (Period) %>%
         filter (Sex==2 & Mat>=stage.mat) %>%  
         summarise(nMatFem= n())

# Females in Mat=4
nSpawFem <- adults.dat %>% group_by (Period) %>%
         filter (Sex==2 & Mat==4) %>%  
         summarise(nSpawFem= n())
  

# Females Collected
nMatFem_C <- adults.dat %>% group_by (Period) %>%
            filter (Cod != "") %>%  
            summarise(nMatFem_C= n())

# Females Screened
nMatFem_H <- adults.dat %>% group_by (Period) %>%
            filter (!is.na(Oocyte_H)) %>%  
            summarise(nMatFem_H= n())

# Females with POFage
nPofFem <-adults.dat %>%  group_by (Period) %>%
filter (POFage != "NA") %>%  
  summarise(nPOFAge= n())

# Females batch fecundity
nNcent <- adults.dat %>%group_by (Period) %>%
  filter (Ncent != "NA") %>%  
  summarise(nNcent= n())

# Create a summary table 
left_join(nHaul, nfish, by="Period") %>%
   left_join(., nMatFem, by="Period") %>%
      left_join(., nSpawFem, by="Period")%>%
        left_join(., nMatFem_C, by="Period")%>%
          left_join(., nMatFem_H, by="Period")%>%
            left_join(., nPofFem, by="Period")%>%
              left_join(., nNcent, by="Period")%>%
                relocate (nFish, .before = D_nFish)%>%
                relocate (R_nFish, .before = D_nFish)%>%
                adorn_totals("row")%>%
                replace(is.na(.), 0)%>%
  kbl(caption = "Sampled by period.") %>%
            kable_classic(#full_width = F, 
              html_font = "Cambria")%>%
            column_spec(8:11, background = "lightgrey")
```

```{r echo=T}
# ---- Legend ------ #

# nHaul:	 No. of hauls.
# nFish:	 No. of individual fish measured.
# R_nFish:	 No. of individual fish randomly sampled and measured.
# D_nFish:   No. of individual fish sampled by direct sampling and measured.
# nMatFem:   No. of mature females. 
# nSpawFem:  No. of mature females in maturity stage 4 (Hydrated).
# nMatFem_C: No. of ovaries from mature females that were collected.
# nMatFem_H: No. of ovaries from collected mature females that were histologically screened.
# nPOFAge:   No. of ovaries assigned as "with presence of POF" in histological screening.
# nNcent:    No. of ovaries analysed for batch fecundity.

# ------------------ #

```


**By haul**

```{r echo=F, message=FALSE, warning=FALSE}
# Number of fish
nFish <- adults.dat %>% group_by (Haul) %>% summarise(nFish=n())

# Number of Females
nFemales <- adults.dat %>% group_by (Haul) %>% 
  filter (Sex==2) %>% summarise(nFemales=n())

# Females in Mat>=stage.mat 
nMatFem <- adults.dat %>%  group_by (Haul) %>%
  filter (Sex==2 & Mat>=stage.mat) %>%  
  summarise(nMatFem= n())

# Spawning Females: Females in Mat=4
nSpawFem <- adults.dat %>% group_by (Haul) %>%
  filter (Sex==2 & Mat==4) %>%  
  summarise(nSpawFem= n())

# Females Collected
nMatFem_C <- adults.dat %>% group_by (Haul) %>%
            filter (Cod != "") %>%  
            summarise(nMatFem_C= n())

# Females Screened
nMatFem_H <- adults.dat %>% group_by (Haul) %>%
            filter (!is.na(Oocyte_H)) %>%  
            summarise(nMatFem_H= n())

# Females with POFage
nPOFageFem <- adults.dat %>%  
  group_by (Haul) %>%
  filter (POFage != "NA") %>%  
  summarise(nPOFAge= n())

# Females batch fecundity
nNcent <- adults.dat%>%  group_by (Haul) %>%
  filter (Ncent != "NA") %>%  
  summarise(nNcent= n())

# add Haul information
dataHaul <- unique(data.frame(adults.dat$Period, adults.dat$Haul,
                              adults.dat$Long,adults.dat$Lat,adults.dat$Country))


names(dataHaul) <- c("Period", "Haul", "Long", "Lat", "country")

# Create a summary table 
left_join(dataHaul,nFish, by="Haul") %>%
left_join(.,nFemales, by="Haul") %>%
   left_join(., nMatFem, by="Haul") %>%
      left_join(., nSpawFem, by="Haul")%>%
          left_join(., nMatFem_C, by="Haul")%>%
            left_join(., nMatFem_H, by="Haul")%>%
              left_join(., nPOFageFem, by="Haul")%>%
                left_join(., nNcent, by="Haul")%>%
  adorn_totals("row")%>%
  replace(is.na(.), 0)%>%
  arrange(Period)%>%
  kbl(caption = "Samples by Haul.") %>%
            kable_classic(#full_width = F, 
            html_font = "Cambria")%>%
            column_spec(10:13, background = "lightgrey")
        #row_spec(c(4:9,11:13, 19:21, 24:26, 29, 34), background = "lightgrey")
```


## Plots

**Geographical distribution of nFish**

```{r echo=F, message=FALSE, warning=FALSE}

ggplot(data = world, frame = year) +
  geom_sf(color = "black", fill = "lightgrey") +
  coord_sf(xlim = c(-20,12), ylim = c(35,64), expand = T, crs="+proj=longlat   +ellps=WGS84 +datum=WGS84") +
  geom_point(data = adults.dat%>%dplyr::select(Lat, Long, Period)%>%
                group_by(Lat)%>%
                dplyr::mutate(number=n())%>%distinct(), 
                aes(x=Long , y=Lat, size=sqrt(number)*100, fill=Period),
                #aes(x=StartLongitude , y= StartLatitude, size=sqrt(number)*100), fill="darkblue",
                shape = 21, alpha=.5)+
  #scale_size_area(limits=c(0,1000), labels = c("1", "5", "10", "25"),breaks = c( 100,500, 700,1000))+
  theme_bw()+
  labs(size="sample size", title="Nfish samples")+
  theme( legend.position = "right",
         legend.text = element_text(size = 12, colour = "black"),
         legend.title = element_text(face = "bold")
         #legend.box.background = element_rect(color="black", size=1)
         )

```


**Length and weight**

```{r echo=F, message=FALSE, warning=FALSE}
#by period, length
ggplot(adults.dat, aes(x=Length, group=Period)) + 
  geom_histogram(aes(y=..density..), bins = 30, col="black", alpha=0.3, fill="black")+
  geom_density(lwd=1)+
  facet_wrap(~Period)+
  xlab("Fish Length (cm)")

#by period, weight
ggplot(adults.dat, aes(x=Wt, group=Period)) + 
  geom_histogram(aes(y=..density..), bins = 30, col="black", alpha=0.3, fill="grey")+
  geom_density(lwd=1)+
  facet_wrap(~Period)+
  xlab("Fish weight (g)")

#by period, ovaryweight
ggplot(adults.dat, aes(x=Wov, group=Period)) + 
  geom_histogram(aes(y=..density..), bins = 30, col="black", alpha=0.3, fill="white")+
  geom_density(lwd=1)+
  facet_wrap(~Period)+
  xlab("Ovary weigth (g)")
```

# Paremeters by haul and Period {.tabset .tabset-tabs}

```{r echo=T}
# Wh:	Mean weight of mature females
# Rh:	Sex-ratio 
# Fh:	Batch fecundity (the mean female expected batch fecundity)
# Sh:	Spawning fraction ratio 
```

## Total Weight correction {.tabset .tabset-tabs}

```{r echo=F, message=FALSE, warning=FALSE}

# Correct bias produced by hydrated females in expected weight

# Empirical weight-length plot for all females
plot(adults.dat$Length[adults.dat$Sex==2],adults.dat$Wt[adults.dat$Sex==2],
     xlab="Length (cm)", ylab="Total weight (g)", 
     main="Females: Empirical Total Weight-Length relationship", col="red")

# remove outlier <- no lo hago
# adults.dat[adults.dat$Wt > 200 & adults.dat$Length,]$Wt <- NA
#Plot wit outlier corrected
#plot(adults.dat$Length[adults.dat$Sex==2],adults.dat$Wt[adults.dat$Sex==2],xlab="Length (mm)", #ylab="Total weight (g)", main="Females: Empirical Total Weight-Length", col="red")

# Correction:
# Select females, adult (Mat >= 2) and non-hydrated (Mat!=4)
adults.dat$Wnov<- adults.dat$Wt-adults.dat$Wov
adults.dat$Wnov[is.na(adults.dat$Wov)] <- NA
which.weight <-  subset (adults.dat, Sex==2 & Mat>=stage.mat & Mat!=4)

# Fit linear regression to Wt female and Wnov
correctW.lm <- lm(Wt ~ Wnov, data=which.weight)

# Examining the fit 
sum_lm <-summary(correctW.lm)
#coef(correctW.lm)[c(1,2)]

ggplot(which.weight, aes(Wnov, Wt))+
  geom_point()+
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth")
```

```{r echo=T}
# (Intercept)        Wnov 
#  7.795266         1.033756 
#  Adjusted R-squared:  0.99 
```

```{r echo=F, message=FALSE, warning=FALSE}
# Residual Analysis
#plot_res <- plot(correctW.lm)

# Identify outliers
res.dat <- data.frame(correctW.lm$residuals, correctW.lm$fitted.values)
out.lm <- rownames(res.dat[res.dat[,1] > 33.5,])
out.lm <- adults.dat[rownames(adults.dat) %in% out.lm,] 

# Plot: relationship between gonad-free female weight and
# total  weight. Red shaded area = 95% CI
# jpeg (file ="CorrectWeightFemales.jpeg", width = 15, height = 15, units = "cm", res=300)
# par(mfrow=c(1,1))

new <- data.frame (Wnov = seq(min(which.weight$Wnov, na.rm=T), 
                              max(which.weight$Wnov, na.rm=T), len=1000)) 

pred <- predict(correctW.lm, new, type="response", interval="predict")

fit.dat <- cbind(new, pred)

ggplot(fit.dat, aes(Wnov, fit))+
  geom_smooth(method = "lm", se = FALSE)+
  geom_ribbon(aes(ymin = lwr, ymax = upr, color = "red", fill = "red"), alpha = .15)+
  labs(x = "Ovary-free weight (g)", y = "Total weight (g) (Exp)", title = "Correction for hydration.")+
  guides(colour = F, fill = F)+
  geom_point (data =which.weight, aes(Wnov, Wt), col = "navy", pch = 21)

# Prediction (Wexp)!!
# Add values to variable Wexp: expected Weight for Females (Sex==2), Mature (Mat>=stage.mat) 
# from random sampling (Sampling=="A")

adults.dat$Wexp <- ifelse(adults.dat$Sampling=="R" & 
                          adults.dat$Sex==2 & adults.dat$Mat>=stage.mat, 
                          predict(correctW.lm, adults.dat, type="response"), NA)

```


## Female mean weight (Wh)

```{r echo=F, message=FALSE, warning=FALSE}
# females,mature and sampling:random, records with Wexp data:
weight.by.haul <- tapply(adults.dat$Wexp, adults.dat$Haul, mean, na.rm=T)
Wh <- data.frame(Haul=rownames(weight.by.haul), Wh=round(weight.by.haul,3)) 
rownames(Wh) <- NULL
Wh$Wh[is.nan(Wh$Wh)] <- NA

# Number of mature females from random sampling with Wexp data:
females.by.haul <- tapply(adults.dat$Wexp[adults.dat$Sampling=="R"& adults.dat$Sex==2 & adults.dat$Mat>=stage.mat], adults.dat$Haul[adults.dat$Sampling=="R"& adults.dat$Sex==2 & adults.dat$Mat>=stage.mat], length)
females.by.haul[is.na(females.by.haul)] <- 0
nF <- data.frame(Haul=rownames(females.by.haul), n.Females.R=females.by.haul)
rownames(nF) <- NULL

# Wh results
# All.data
n <- data.frame(adults.dat %>%group_by(Period,Haul)%>%dplyr::summarise(n()))
names(n) <- c("Period","Haul", "N")
nMales <- data.frame(adults.dat %>% filter(Sex==1 & Mat>=stage.mat) %>%group_by(Period,Haul)%>%dplyr::summarise(n()))
names(nMales) <- c("Period","Haul", "nMales")
nFemales <- data.frame(adults.dat %>% 
                         filter(Sex==2 & Mat>=stage.mat) %>%group_by(Period,Haul)%>%dplyr::summarise(n()))
names(nFemales) <- c("Period", "Haul", "nFemales")
NbyHaul <- merge( merge(merge (n, nMales,all=T), nFemales,all=T), nF, all=T)
NbyHaul[is.na(NbyHaul)] <- 0

# minimum number of females to estimate mean weight by haul
Wh <- merge(NbyHaul, Wh, by="Haul")
Wh$Wh <- ifelse (Wh$n.Females.R < min.totnumb, NA, Wh$Wh)

Wh%>%relocate(Period, .before = Haul) %>%
  arrange(Period)%>%
    kbl(caption = "Parameters: Wh by haul (auxiliar).") %>%
            kable_classic(#full_width = F, 
            html_font = "Cambria")
```


## Sex ratio (Rh)

```{r echo=F, message=FALSE, warning=FALSE}

# Mean weight of adult males from random sampling with Wt data
malew.by.haul <- tapply(adults.dat$Wt[adults.dat$Sampling=="R"& adults.dat$Sex==1 & adults.dat$Mat>=stage.mat],adults.dat$Haul[adults.dat$Sampling=="R"& adults.dat$Sex==1 & adults.dat$Mat>=stage.mat], mean, na.rm=T)
malew.by.haul[is.na(malew.by.haul)] <- 0 # remove NA, it is neccesary to sum weight.females+weight.males
malew.by.haul <- as.data.frame(malew.by.haul)

# Number of adult males from random sampling with Wt data
malen.by.haul <- tapply(adults.dat$Wt[adults.dat$Sampling=="R"& adults.dat$Sex==1 & adults.dat$Mat>=stage.mat],adults.dat$Haul[adults.dat$Sampling=="R"& adults.dat$Sex==1 & adults.dat$Mat>=stage.mat], length)
malen.by.haul[is.na(malen.by.haul)] <- 0 # remove NA
malen.by.haul <- as.data.frame(malen.by.haul)

totnumb.by.haul <- data.frame(females.by.haul + malen.by.haul)
names(totnumb.by.haul) <- "TotNh"
totweight.by.haul <- data.frame(weight.by.haul + malew.by.haul)
names(totweight.by.haul) <- "TotWh"
ratio.by.haul <- data.frame(weight.by.haul/ totweight.by.haul)
names(ratio.by.haul) <- "Rh"

Rh <- data.frame(rownames(totnumb.by.haul), totnumb.by.haul, round(ratio.by.haul,3))
names(Rh) <- c("Haul", "TotNh", "Rh")
rownames(Rh) <- NULL

# Hauls with numbers < min.totnumb, remove values of ratio 
data.H <- merge (Wh, Rh, by="Haul", all=TRUE)
data.H$Rh <- ifelse (data.H$TotNh < min.totnumb, NA, data.H$Rh)

data.H%>%relocate(Period, .before = Haul) %>%
  arrange(Period)%>%
  kbl(caption = "Parameters: Rh by haul (auxiliar).") %>%
            kable_classic(#full_width = F, 
            html_font = "Cambria")
```


## Batch fecundity (Fh) 

**Geographical distribution_nCent**

```{r echo=F, message=FALSE, warning=FALSE}

ggplot(data = world, frame = year) +
  geom_sf(color = "black", fill = "lightgrey") +
  coord_sf(xlim = c(-20,12), ylim = c(35,64), expand = T, crs="+proj=longlat   +ellps=WGS84 +datum=WGS84") +
  geom_point(data = adults.dat%>%
                group_by(Lat)%>%
                dplyr::mutate(number=n()), 
                aes(x=Long , y= Lat, 
                    size=number, fill= Period),
                #aes(x=StartLongitude , y= StartLatitude, size=sqrt(number)*100), fill="darkblue",
                shape = 21, alpha=.9)+
  scale_size_area(limits=c(0,500), labels = c("1", "5", "10", "25"),breaks = c( 100,224, 320,500))+
  theme_bw()+
  labs(size="sample size", title="Nfish samples")+
  theme( legend.position = "right",
         legend.text = element_text(size = 12, colour = "black"),
         legend.title = element_text(face = "bold")
         #legend.box.background = element_rect(color="black", size=1)
         )

```



**Fit regression**

```{r echo=F, message=FALSE, warning=FALSE}

# Fit a lm, glm, gaussian, removing records with Fobs values > 41940
adults.dat <- adults.dat%>%mutate(batch.ov = ifelse(!is.na(Ncent), round((Wov*Ncent)/Wcent, 0),NA))

aux <- ggplot(adults.dat, aes(x = 1, y = batch.ov))+
          geom_point(colour = "red", size = 8, shape = 21)+
          theme(axis.text.x=element_blank(),
          axis.ticks.x=element_blank())

# the individual observed batch fecundities (Fobs) are first
# modelled against the corresponding female gonad-free weights
# linear. it is assumed that the relationship is linear.

batch.lm<- lm(batch.ov ~ Wnov, data=adults.dat, na.action="na.omit")
sum.lm.batch <- summary(batch.lm)
# coef(batch.lm)[c(1,2)]
# y = batch.lm$coefficients[1] + batch.lm$coefficients[2]*x
# R^2 = sqrt(0.1769)

batch.glm.nb <- glm.nb(batch.ov~ Wnov, data=adults.dat,na.action="na.omit")
summary(batch.glm.nb)

finalbatch.glm <- batch.glm.nb

# Plot of fitted model
# data used for the fit
fec.fit.data <- na.omit(adults.dat[, c("Wnov","batch.ov")])
jpeg (file ="PlotFittedModelFecundity.jpeg", width = 15, height = 15, units = "cm", res=300)
par(mfrow=c(1,1))
plot(batch.ov ~ Wnov, data=fec.fit.data, pch=20,xlab="Ovary-free weight (g) ", ylab= "Fecundity (oocytes per batch)", main="Batch Fecundity",xlim=c(150, 500), ylim=c(0,41000))
# Fitted function 
NB.func <- function(x) exp(coef(finalbatch.glm)[1]+ coef(finalbatch.glm)[2]*x)
curve(NB.func, add=T, lty=1, col=2)
legend('topleft', c('negative binomial'), col=c(2), lty=c(1), bty='n', cex=.9)
dev.off()


```

```{r echo=F, message=FALSE, warning=FALSE}
# This model is subsequently applied to all
# mature females in the sample to obtain the expected individual batch fecundity (Fexp), i.e. the
# batch fecundity that these sampled females (with a given body weight) would have if they were
# hydrated at the time of capture.


# Fitted model on all mature females
# fit <- function(x) batch.lm$coefficients[1] + batch.lm$coefficients[2]*x


# Estimate expected Fecundity (Fexp)
# New empty variable, Fexp
adults.dat$Fexp <- NA

# Predict values of Fexp.lab for model choosed:
adults.dat$Fexp[adults.dat$Sampling=="R" & adults.dat$Mat >= stage.mat & !is.na(adults.dat$Sampling=="R" & adults.dat$Mat >= stage.mat)] <- predict(finalbatch.glm, adults.dat[adults.dat$Sampling=="R" & adults.dat$Mat >= stage.mat & !is.na(adults.dat$Sampling=="R" & adults.dat$Mat >= stage.mat),],type="response")

# Batch Fecundity by haul 
batch.by.haul <- tapply(adults.dat$Fexp, adults.dat$Haul, mean, na.rm=T)

Fh <- data.frame(rownames(batch.by.haul), round(batch.by.haul,3))
rownames(Fh) <- NULL
names(Fh) <- c("Haul", "Fh")
Datah2 <- merge(data.H, Fh, by="Haul",all=TRUE)
Datah2$Fh <- ifelse (Datah2$TotNh  < min.totnumb, NA, Datah2$Fh)
Datah2

```

## Spawning fraction (Sh)

```{r echo=F, message=FALSE, warning=FALSE}
# Only for Sampling="A" and Mat > mat.stage 
# numeric variable 
adults.dat$POFage <- as.numeric(as.character(adults.dat$POFage))

table(adults.dat$POFage)

# -9   0   1   2   3 
# 161 354 187 107 124 

# IMPORTANT: Selected data: mature, aleatory sampling and with POFs value (-9 (no POFs),0,1,2,3)
mature   <- adults.dat$Mat >= stage.mat
aleatory <- adults.dat$Sampling=="R" 
POFs <- !is.na(adults.dat$POFage) 
which.sfract <- mature & aleatory & POFs

# Function to estimate Spawning fraction
# Only non-hydrated records (SISP5 recommendation)
# noPOFs (-9 values) are also used in the formula
sfraction.fun <- function(i,data){
  data <- data[data$Haul==i,]
  data <- data[data$Mat!= 4,] # hydrated are not used to estimate Sh
  if(is.na(sum(data$POFage, na.rm=T)))
    sfract <- NA
  else{
    day1day2 <- sum(!is.na(data$POFage[data$POFage==1 | data$POFage==2]))
    day3plus <- sum(!is.na(data$POFage[data$POFage>2]))
    day3plus <- day3plus + sum(!is.na(data$POFage[data$POFage == -9]))
    sfract <- (day1day2/2)/((day1day2/2)+day1day2+day3plus)
  }
  sfract
}

sfract.by.haul <- sapply(unique(adults.dat$Haul[which.sfract]), 
                         sfraction.fun, adults.dat[which.sfract,])

sfract.by.haul <- sfract.by.haul[match(names(females.by.haul), 
                                       unique(adults.dat$Haul[aleatory]))]

names(sfract.by.haul) <- names(females.by.haul)

Sh <- data.frame(names(sfract.by.haul), round(sfract.by.haul,3))
rownames(Sh) <- NULL
names(Sh) <- c("Haul", "Sh")
data.SH <- merge(Datah2, Sh, by="Haul",all=TRUE)
data.SH$Sh <- ifelse (data.SH$TotNh < min.totnumb, NA, data.SH$Sh)

data.SH%>%relocate(Period, .before = Haul) %>%
  arrange(Period)%>%
        kbl(caption = "Parameters by haul.") %>%
            kable_classic(#full_width = F, 
            html_font = "Cambria")

```


# Population parameters by Period {.tabset .tabset-tabs}

```{r echo=T, message=FALSE, warning=FALSE}
# W:  Mean weight of mature females
# R:  Sex-ratio
# F:  Batch fecundity = average number of egg by mature female by spawning event (g)
# S:  Spawning fraction 
```

```{r echo=F, message=FALSE, warning=FALSE}

# Minimum samples
min.totnumb <- 20
totnumb.by.haul.c <- totnumb.by.haul
totnumb.by.haul.c[totnumb.by.haul < min.totnumb] <- NA

# Function to estimate variance following the methodology from Picquelle and Stauffer (1985):
var.DEPM <- function(data,wt)
  {
  DATpop <- wt.mean(data, wt)
  n <- length(data)
  return(sum(wt^2*(data - DATpop)^2, na.rm=T) / (sum(wt/n,na.rm=T)^2*n*(n-1)))
  }
```

## Mean Weight

```{r echo=F, message=FALSE, warning=FALSE}

Wmean <-data.SH%>%filter(!is.na(Wh))%>%
            summarise(mean = wt.mean(Wh,n.Females.R),
                      var = var.DEPM(Wh,n.Females.R), 
                    cv = sqrt(var/mean))

WmeanP <-data.SH%>%filter(!is.na(Wh))%>%
            group_by(Period)%>%
            summarise(mean = wt.mean(Wh,n.Females.R),
                      var = var.DEPM(Wh,n.Females.R), 
                    cv = sqrt(var/mean))
Wmean
WmeanP
```                  

## Mean Sex Ratio                 

```{r echo=F, message=FALSE, warning=FALSE}

Rmean <-data.SH%>%filter(!is.na(Rh))%>%
            summarise(mean = wt.mean(Rh,TotNh),
                      var = var.DEPM(Rh,TotNh), 
                      cv = sqrt(var/mean))

RmeanP <-data.SH%>%filter(!is.na(Rh))%>%
          group_by(Period)%>%
            summarise(mean = wt.mean(Rh,TotNh),
                      var = var.DEPM(Rh,TotNh), 
                      cv = sqrt(var/mean))
Rmean
RmeanP
```

## Mean Batch Fecundity
    
```{r echo=F, message=FALSE, warning=FALSE}

Fmean <-data.SH%>%filter(!is.na(Fh))%>%
            summarise(mean = wt.mean(Fh,n.Females.R),
                      var = var.DEPM(Fh,n.Females.R), 
                    cv = sqrt(var/mean))

FmeanP <-data.SH%>%filter(!is.na(Fh))%>%
            group_by(Period)%>%
            summarise(mean = wt.mean(Fh,n.Females.R),
                      var = var.DEPM(Fh,n.Females.R), 
                    cv = sqrt(var/mean))
Fmean
FmeanP

```

## Mean Spawning fraction
 
```{r echo=F, message=FALSE, warning=FALSE}
Smean <-data.SH%>%filter(!is.na(Sh))%>%
            summarise(mean = wt.mean(Sh,n.Females.R),
                      var = var.DEPM(Sh,n.Females.R), 
                      cv = sqrt(var/mean))

SmeanP <-data.SH%>%filter(!is.na(Sh))%>%
          group_by(Period)%>%
            summarise(mean = wt.mean(Sh,n.Females.R),
                      var = var.DEPM(Sh,n.Females.R), 
                      cv = sqrt(var/mean))

Smean 
SmeanP 
```

# Summary mean parameters

```{r echo=F, message=FALSE, warning=FALSE}

options(scipen=999)

bind_rows (Wmean, Rmean, Smean, Fmean)%>%
  mutate(Parameter=c("Female Mean weight ", "Mean Sex Ratio", "Mean Spawning Fraction", "Mean batch fecundity"))%>%
  relocate(Parameter, .before = mean)%>%
  add_row(Parameter = "Mean batch fecundity/g Female (Frel)", 
          mean = 9374.3285417/355.9464902 , var = NA, cv = NA)%>%
  add_row(Parameter = "Daily n.eggs/g Female", 
          mean = 9374.3285417/355.9464902 * 0.5096771* 0.1504741 , var = NA, cv = NA)%>%
  mutate(across(2:4, round, 3))%>%# 29 days
  kbl(caption = "Adult parameters estimates for the 2022 Western Atlantic mackerel DEPM survey.") %>%
            kable_classic(#full_width = F, 
            html_font = "Cambria")
```



# Summary SSB

```{r echo=F, message=FALSE, warning=FALSE}

# ssb <- adults.dat <- read.csv("dat/HOM_PortalSSB.csv")
# names(ssb) <- c("Period", "Daily fecundity", "Total daily egg production(e+12)", "SSB(kt)")

#ssb%>%kbl(caption = "SSB by period based on aduts parameters by period and combined.") %>%
            # kable_classic(#full_width = F, 
            # html_font = "Cambria")

```


```{r echo=T}
# @ https://github.com/GersomCostas/mackerel_adults_MEGS
# @ https://github.com/ices-eg/wg_WGMEGS 
```
